Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY: Supplementary materials

### 2022-03-03 manuscript/draft-supplement.Rmd
lots of updates to manuscript

# where are the outputs (ie the inputs for this manuscript!) saved?
output_dir_os <- here("output", "manuscript-objects")
#output_dir_os <- here("released_outputs", "output", "manuscript-objects")

@@

flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
smd <- read_csv(fs::path(output_dir_os, "matchsmd.csv"))

@@

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath"
    ) %>% 
    set_names(., .)

@@

km_none <- read_csv(fs::path(output_dir_os, "none", "km.csv"))  
km_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")) 
km <- 
  bind_rows(km_none, km_vax12_type) %>%
  mutate(
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
  )

hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv"))
hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv"))
hazards <- 
  bind_rows(hazards_none, hazards_vax12_type) %>%
  mutate(

@@

Note that we excluded care home residents and health care workers in our analysis, so group 1 is not included and group 2 includes only those aged 80 and over. The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else". Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligiblity in the 40-49 group from 15 November 2021 onwards. (https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)

@@

plot_smd <-
  ggplot(data_smd)+
  geom_point(aes(x=smd, y=level))+
  geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5)), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
  scale_alpha_continuous(range=c(0,0.3), guide=FALSE)+
  labs(
    x="Standardised mean difference",
    y=NULL,
    alpha=NULL
  )+
  theme_minimal() +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(fill="transparent", colour="transparent"),
    strip.text.y.left = element_text(angle = 0, hjust=1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.spacing = unit(0, "lines")
  )

@@

plot_km <- km %>%
    ggplot()+
    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
    geom_hline(aes(yintercept=0), colour='black')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
    scale_x_continuous(
      breaks = seq(0,7*max(postbaselinecuts),by=14),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0),
      #limits = c(0L,NA)
    )+
    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
    labs(
      x=NULL,
      y="Cumulative incidence\nper 1000",
      colour=NULL
    )+
    theme_minimal()+
    theme(
      legend.position=c(0.15,0.85),
      legend.justification = c(1,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

  plot_hazards <-
    hazards %>%
    filter(
      outcome %in% outcomes,
      treatment %in% c("pfizer")
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

>>>>

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "manuscript-objects")
output_dir_os <- here("released_output",  "manuscript-objects")

@@

matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
smd <- read_csv(fs::path(output_dir_os, "matchsmd.csv"))

@@

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

@@

km <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "km.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")) ,
    #read_csv(fs::path(output_dir_os, "cev", "km.csv")) ,
    #read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) ,
  ) %>%
  mutate(
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
  )

hazards <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    #read_csv(fs::path(output_dir_os, "cev", "effects.csv")),
    #read_csv(fs::path(output_dir_os, "age65plus", "effects.csv")),
  ) %>%
  mutate(

@@

Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over. The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else". Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligiblity in the 40-49 group from 15 November 2021 onwards. (https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)

# Variables and codelists

```{r, codelists, echo=FALSE, message=FALSE, warning=FALSE}

variable_table <- readxl::read_xlsx(here("lib", "design", "variable-descriptions.xlsx"))

variable_table %>%
  gt(groupname_col = "category") %>%
  cols_label(
    name = "Variable",
    source = "Data source",
    description = "Notes",
    support = "Values ",
    codelistID = "Codelist identifier"
  ) %>%
   fmt_missing(
    everything(),
    missing_text=""
  )

Codelists can be found at `https://codelists.opensafely.org/codelist/<ID>`, substituting `<ID>` for the codelist identifier in the table above.

@@

plot_smd <-
  smd %>%
  mutate(
    level=fct_rev(fct_inorder(level))
  ) %>%
  ggplot()+
  geom_point(aes(x=smd, y=level))+
  geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5)), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
  scale_alpha_continuous(range=c(0,0.3), guide=FALSE)+
  labs(
    x="Standardised mean difference",
    y=NULL,
    alpha=NULL
  )+
  theme_minimal() +
  theme(
    strip.placement = "outside",
    strip.background = element_rect(fill="transparent", colour="transparent"),
    strip.text.y.left = element_text(angle = 0, hjust=1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.spacing = unit(0, "lines")
  )

@@

plot_km <- km %>%
  filter(
    outcome %in% outcomes,
    treatment=="pfizer"
  ) %>%
    ggplot()+
    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
    geom_hline(aes(yintercept=0), colour='black')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
    scale_x_continuous(
      breaks = seq(0,7*max(postbaselinecuts),by=14),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0),
      #limits = c(0L,NA)
    )+
    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
    labs(
      x=NULL,
      y="Cumulative incidence\nper 1000",
      colour=NULL
    )+
    theme_minimal()+
    theme(
      legend.position=c(0.15,0.85),
      legend.justification = c(1,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

@@

  plot_hazards <-
    hazards %>%
    filter(
      outcome %in% outcomes,
      treatment %in% c("pfizer")
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 4))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),
